\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2018-11-07 Documents/M1/Astronum_2018/sections/TimeIntegration.tex
Modified time integration section.

Suppose that an algebraic closure based on Fermi-Dirac statistics is used (i.e., the Eddington factor satisfies Eq.~\eqref{eq:eddingtonFactorBounds}).
Here we consider the construction of an Implicit-Explicit (IMEX) time integration scheme which maintains the bounds in Eq.~\eqref{eq:MomentsBounds}.  
The semi-discretization of the two-moment model results in a system of ordinary differential equations of the form
\begin{equation}
  \dfrac{d \vect{u}}{d t} = \vect{\cT}(\vect{u}) + \f{1}{\tau}\,\vect{\cQ}(\vect{u}),
\end{equation}
where the solution vector
\begin{equation}
  \vect{u}(t) = \left( \bcM_{1}(t),\ldots,\bcM_{N}(t)\right) ^{T}
\end{equation}
is the collection of all cell-averaged moments, $\vect{\cT}$ is the transport operator, corresponding to the first term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}, and $\f{1}{\tau}\,\vect{\cQ}$ is the collision operator, corresponding to the second term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}.

@@

Treating the transport operator explicitly and the collision operator implicitly, a standard $s$-stage IMEX scheme takes the following form \cite{pareschiRusso_2005}: 
\begin{align}
  \vect{u}^{(i)}
  &=\vect{u}^{n}
  +\dt\sum_{j=1}^{i-1}\tilde{a}_{ij}\,\vect{\cT}(\vect{u}^{(j)})
  +\dt\sum_{j=1}^{i}a_{ij}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(j)}),
  \quad i=1,\ldots,s, \label{imexStages} \\
  \vect{u}^{n+1}
  &=\vect{u}^{n}
  +\dt\sum_{i=1}^{s}\tilde{w}_{i}\,\vect{\cT}(\vect{u}^{(i)})
  +\dt\sum_{i=1}^{s}w_{i}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}), \label{imexIntermediate} 
\end{align}
where $(\tilde{a}_{ij})$ and $(a_{ij})$, coefficients of the $i$-th stage, are the elements of matrices $\tilde{A}$ and $A$, respectively.
The vectors $\tilde{\vect{w}}=(\tilde{w}_{1},\ldots,\tilde{w}_{s})^{T}$ and $\vect{w}=(w_{1},\ldots,w_{s})^{T}$ are the weights in the assembly step in Eq.~\eqref{imexIntermediate}.
These coefficients and weights must satisfy certain order conditions for consistency and accuracy.  
For second-order temporal accuracy, the following conditions are required:
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=\sum_{i=1}^{s}w_{i}=1,
  \label{orderConditions1}
\end{equation}
and
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}\tilde{w}_{i}\,c_{i}
  =\sum_{i=1}^{s}w_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}w_{i}\,c_{i}=\f{1}{2}, 
  \label{orderConditions2}
\end{equation}
where $\tilde{c}_{i} = \sum_{j=1}^{s}\tilde{a}_{ij}$ and $c_{i}=\sum_{j=1}^{s}a_{ij}$.

\ee{Discuss GSA and Type~ARS here.}

@@

Following Hu et al\cite{hu_etal_2018}, the stage values in Eq.~\eqref{imexStages} can be rewritten as
\begin{equation}
  \vect{u}^{(i)}
  =\sum_{j=0}^{i-1}c_{ij}\Big[\,\vect{u}^{(j)}+\hat{c}_{ij}\,\dt\,\vect{\cT}(\vect{u}^{(j)})\,\Big]
  +a_{ii}\,\dt\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}),\quad i=1,\ldots,s,
  \label{eq:imexStagesRewrite}
\end{equation}
with $c_{ij}$ and $\hat{c}_{ij}$ be some expression of $a_{ij}$ and $\tilde{a}_{ij}$.
By requiring the IMEX scheme be globally stiffly accurate (GSA), i.e. $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i}$ for $i=1,\ldots,s$, $\vect{u}^{n+1} = \vect{u}^{(s)}$.
Using those Lemma proved in \cite{chu_etal_2018}, we can find those constraints that $c_{ij}$ and $\hat{c}_{ij}$ (or $a_{ij}$ and $\tilde{a}_{ij}$) need to satisfy to ensure $\vect{u}^{(i)}$ and $\vect{u}^{n+1}$ be realizable.

\ee{Move definition of coefficients $c_{ij}$ and $\tilde{c}_{ij}$ here}

\subsection{Diffusion Accurate, Constraint-Preserving Implicit-Explicit Schemes}
Diffusion accurate is the other property the time integration should have.
In the diffusion limit, the distribution function is nearly isotropic so that $\vect{\cK}\approx\f{1}{3}\,\cJ\,\vect{I}$ and $\vect{\cH}\approx-\f{1}{3}\,\tau\,\nabla\cJ$.
The moment behavior are governed by (e.g., \cite{jinLevermore_1996})
\begin{equation}
  \pd{\cJ}{t} + \nabla\cdot\vect{\cH} = 0
  \quad\text{and}\quad
  \vect{\cH} = - \tau\,\nabla\cdot\vect{\cK}.  
  \label{eq:diffusionLimit}
\end{equation}
In the context of IMEX schemes, the above relationships imply that the relation $A\,\vec{\vect{\cH}}=-\f{1}{3}\,\tau\,\tilde{A}\,\nabla\vec{\cJ}$ should hold. 
We derived the diffusion accurate condition in \cite{chu_etal_2018} and proved that only type~ARS (investigated by Ascher, Ruuth, and Spiteri\cite{ascher_etal_1997}) IMEX schemes can be our candidates.

For type~ARS, $a_{i1} = 0$ for $i=1,\ldots,s$.
$c_{ij}$ and $\hat{c}_{ij}$ are given by
    \begin{equation}
     \begin{aligned}
      c_{i0} &= 1-\sum_{j=2}^{i-1}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad &
      c_{ij} &= \sum_{l=j}^{i-1}a_{il}b_{lj}, \\
      \tilde{c}_{i0} &= \tilde{a}_{i1}+\sum_{j=2}^{i-1}a_{ij}\tilde{b}_{j1}, \quad &
      \tilde{c}_{ij} &= \tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj},  
     \end{aligned}
     \label{eq:positivityCoefficientsARS}
    \end{equation}
    \begin{equation}
      b_{ii} = \f{1}{a_{ii}}, \quad
      b_{ij} = -\f{1}{a_{ii}}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad
      \tilde{b}_{ij} = -\f{1}{a_{ii}}\Big(\tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}\Big).  
    \end{equation}
    Note that $c_{i1}=\tilde{c}_{i1}=0$ in Eq.~\eqref{eq:positivityCoefficientsARS} so that $\sum_{j=0}^{i-1}c_{ij}=1$.

Combining the conditions we have, we aim to find type~ARS IMEX scheme with.
\begin{enumerate}
    \item Consistency of the implicit coefficients:
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
    \end{equation}
    \item Second-order accuracy in the streaming limit:
    \begin{equation}
      \sum_{i=1}^{s}\tilde{w}_{i}=1
      \quad\text{and}\quad
      \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},
      \label{eq:orderConditionsEx}
    \end{equation}
    \item Convex-invariant:
    \begin{align}
      &a_{ii}>0, \quad c_{i0},\tilde{c}_{i0}\ge0, \quad \text{for} \quad i=2,\ldots,s, \nonumber \\
      &\text{and} \quad c_{ij},\tilde{c}_{ij}\ge0, \quad \text{for} \quad i=3,\ldots,s, \quad\text{and}\quad j=2,\ldots,i-1.  
    \end{align}
    with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\Sch}=\min_{\substack{i = 2,\ldots,s \\ 
                  j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}}>0$.
    \item Well-behaved in the diffusion limit:
    \begin{equation}
      \vect{e}_{i}^{T}A^{-1}\tilde{A}\,\vect{e} = 1, \quad i=2,\ldots,s,
      \label{eq:diffusionCondition}
    \end{equation}
    where $\vect{e}_{i}$ is the $i$th column of the $s\times s$ identity matrix.
    \item Less than five stages ($s\le4$).
    \item Globally stiffly accurate: $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$. 
\end{enumerate}  
We call the IMEX scheme satisfying the above conditions {PD-ARS}. (Definition 3 in \cite{chu_etal_2018}.)
We give two optimized PD-ARS schemes here: PD-ARS with SSPRK2 and  PD-ARS with SSPRK3. 
They have second- and third-order accuracy in the streaming limit, respectively.
\subsubsection{PD-ARS with SSPRK2}
Here we give the optimized PD-ARS with SSPRK2 in the standard double Butcher tableau form: explicit tableau on the left ($\tilde{A}$), implicit tableau on the right ($A$):
\begin{equation}
  \begin{array}{c | c c c}
  	0 & 0   & 0 & 0 \\
  	1 & 1   & 0 & 0 \\
  	1 & 1/2 & 1/2 & 0 \\ \hline
  	  & 1/2 & 1/2 & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1 & 0 & 1/2 & 1/2 \\ \hline
  	  & 0 & 1/2 & 1/2
  \end{array}
\end{equation}
For this scheme, $c_{\mbox{\tiny Sch}}= 1$ and only two implicit solver are needed for each time step.
\subsubsection{PD-ARS with SSPRK3}
Here we give the optimized PD-ARS with SSPRK3.
Its standard double Butcher tableau are (explicit tableau on the left, implicit tableau on the right)
\begin{equation}
  \begin{array}{c | c c c c}
  	    &     &     &     &  \\
  	 1  & 1   &     &     &  \\
  	1/2 & 1/4 & 1/4 &  \\
  	 1  & 1/6 & 1/6 & 2/3 &  \\ \hline
  	    & 1/6 & 1/6 & 2/3 &
  \end{array}
  \qquad
  \begin{array}{c | c c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1/2 & 0 & 1/4 & 1/4 \\ 
  	1 & 0 & 1/6 & 1/6 & 2/3\\\hline
  	  & 0 & 1/6 & 1/6 & 2/3
  \end{array}
\end{equation}
For this schemes, $c_{\mbox{\tiny Sch}}= 1$.
Three implicit solver are taken for each time step.
Since PD-ARS with SSPRK3 is as accurate as PD-ARS with SSPRK2 in collision involved region, see Section~\ref{se:NumericalTests} for the plots, it may not offer any advantage.

\clearpage

>>>>

Suppose that an algebraic closure based on Fermi-Dirac statistics is used (i.e., the Eddington factor satisfies Eq.~\eqref{eq:eddingtonFactorBounds}).
Here we consider the construction of an Implicit-Explicit (IMEX) time integration scheme which maintains the bounds in Eq.~\eqref{eq:MomentsBounds}.  
The semi-discretization of the two-moment model results in a system of ordinary differential equations of the form
\begin{equation}
  \dfrac{d \vect{u}}{d t} = \vect{\cT}(\vect{u}) + \f{ 1}{\tau}\,\vect{\cQ}(\vect{u}),
\end{equation}
where the solution vector
\begin{equation}
  \vect{u}(t) = \left( \bcM_{1}(t),\ldots,\bcM_{N}(t)\right) ^{T}
\end{equation}
is the collection of all cell-averaged moments, $\vect{\cT}$ is the transport operator, corresponding to the first term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}, and $\f{1}{\tau}\,\vect{\cQ}$ is the collision operator, corresponding to the second term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}.

@@

Treating the transport operator explicitly and the collision operator implicitly, a standard $s$-stage IMEX scheme takes the following form \cite{pareschiRusso_2005}: 
\begin{align}
  \vect{u}^{(i)}
  &=\vect{u}^{n}
  +\dt\sum_{j=1}^{i-1}\tilde{a}_{ij}\,\vect{\cT}(\vect{u}^{(j)})
  +\dt\sum_{j=1}^{i}a_{ij}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(j)}),
  \quad i=1,\ldots,s, \label{imexStages} \\
  \vect{u}^{n+1}
  &=\vect{u}^{n}
  +\dt\sum_{i=1}^{s}\tilde{w}_{i}\,\vect{\cT}(\vect{u}^{(i)})
  +\dt\sum_{i=1}^{s}w_{i}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}), \label{imexIntermediate} 
\end{align}
where $(\tilde{a}_{ij})$ and $(a_{ij})$, coefficients of the $i$-th stage, are the elements of matrices $\tilde{A}$ and $A$, respectively.
Matrices $\tilde{A}$ and $A$ are both lower triangular matrices.
The vectors $\tilde{\vect{w}}=(\tilde{w}_{1},\ldots,\tilde{w}_{s})^{T}$ and $\vect{w}=(w_{1},\ldots,w_{s})^{T}$ are the weights in the assembly step in Eq.~\eqref{imexIntermediate}.
These coefficients and weights must satisfy certain order conditions for consistency, accuracy and other properties.

For second-order temporal accuracy, the following conditions are required \cite{hairer_1981}:
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=\sum_{i=1}^{s}w_{i}=1,
  \label{orderConditions1}
\end{equation}
and
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}\tilde{w}_{i}\,c_{i}
  =\sum_{i=1}^{s}w_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}w_{i}\,c_{i}=\f{1}{2}, 
  \label{orderConditions2}
\end{equation}
where $\tilde{c}_{i} = \sum_{j=1}^{s}\tilde{a}_{ij}$ and $c_{i}=\sum_{j=1}^{s}a_{ij}$.

For globally stiffly accuracy (GSA), the coefficients have the following relationship \cite{dimarcoPareschi2013}:
\begin{equation}
a_{si}=w_{i}, \quad \tilde{a}_{si}=\tilde{w}_{i}, \quad \text{for} \quad i=1,\ldots,s.
\end{equation}

The IMEX schemes are usually classified by the structure of matrix $A$, the implicit tableau:
if $A$ is invertible, the IMEX scheme is type~A \cite{pareschiRusso_2005};
if $a_{i1} = 0$ for $i=1,\ldots,s$ and $w_{1} = 0$, the IMEX scheme is type~ARS \cite{ascher_etal_1997,pareschiRusso_2005}.

@@

Following Hu et al\cite{hu_etal_2018}, the stage values in Eq.~\eqref{imexStages} can be rewritten as
\begin{equation}
  \vect{u}^{(i)}
  =\sum_{j=0}^{i-1}c_{ij}\Big[\,\vect{u}^{(j)}+\hat{c}_{ij}\,\dt\,\vect{\cT}(\vect{u}^{(j)})\,\Big]
  +a_{ii}\,\dt\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}),\quad i=1,\ldots,s,
  \label{eq:imexStagesRewrite}
\end{equation}
with $c_{ij}$ and $\hat{c}_{ij}$ be some expression of $a_{ij}$ and $\tilde{a}_{ij}$.
For example, type~ARS IMEX schemes have $c_{ij}$ and $\hat{c}_{ij}$ be given by
    \begin{equation}
     \begin{aligned}
      c_{i0} &= 1-\sum_{j=2}^{i-1}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad &
      c_{ij} &= \sum_{l=j}^{i-1}a_{il}b_{lj}, \\
      \tilde{c}_{i0} &= \tilde{a}_{i1}+\sum_{j=2}^{i-1}a_{ij}\tilde{b}_{j1}, \quad &
      \tilde{c}_{ij} &= \tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj},  
     \end{aligned}
     \label{eq:positivityCoefficientsARS}
    \end{equation}
    \begin{equation}
      b_{ii} = \f{1}{a_{ii}}, \quad
      b_{ij} = -\f{1}{a_{ii}}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad
      \tilde{b}_{ij} = -\f{1}{a_{ii}}\Big(\tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}\Big).  
    \end{equation}
Note that $c_{i1}=\tilde{c}_{i1}=0$ in Eq.~\eqref{eq:positivityCoefficientsARS} so that $\sum_{j=0}^{i-1}c_{ij}=1$.

By requiring the IMEX scheme be GSA, $\vect{u}^{n+1} = \vect{u}^{(s)}$.
Then using those Lemma proved in \cite{chu_etal_2018}, we can find those constraints, such as Eq.~\eqref{eq:convexInvariant} for type~ARS, that $c_{ij}$ and $\hat{c}_{ij}$ (or $a_{ij}$ and $\tilde{a}_{ij}$) need to satisfy to ensure $\vect{u}^{(i)}$, Eq.~\eqref{eq:imexStagesRewrite}, be realizable.
Hence $\vect{u}^{(i)}$ and $\vect{u}^{n+1}$ are realizable and the IMEX is constraint-preserving.

When the convex-invariant constraints are satisfied, the first (sum) term on the right-hand side of Eq.\eqref{eq:imexStagesRewrite} is a convex combination of multiple explicit Euler schemes with time step $\hat{c}_{ij}\,\dt$.
Each of the explicit Euler schemes has a time step condition that ensures its realizability: $\hat{c}_{ij}\,\dt\leq$ the Courant-Friedrichs-Lewy (CFL) condition.
As a result, the time step of the IMEX scheme is subject to the most restrict time step condition:
\begin{equation}
\max(\hat{c}_{ij})\,\dt \leq \text{CFL condition}.
\end{equation}

Unfortunately, no coefficients can be found for standard IMEX schemes under the constraint-preserving constraints and the accuracy constraints for 2nd-order or higher. 
To circumvent this problem, some correction steps are introduced after the assembly step, Eq.~\eqref{imexIntermediate}, such as Hu et al.\cite{hu_etal_2018}.
However, the correction steps impose extra constraints for realizability on time step.
In the circumstance that we have, correction steps can ruin the efficiency we gain from the IMEX scheme.
To keep things simple, we aim for 1st-order in general, high-order in streaming limit, diffusion accurate, constraint-preserving IMEX schemes.

\subsection{Diffusion Accurate, Constraint-Preserving Implicit-Explicit Schemes}

Diffusion accurate is an important property the time integration should have.
In the diffusion limit, the distribution function is nearly isotropic so that $\vect{\cK}\approx\f{1}{3}\,\cJ\,\vect{I}$ and $\vect{\cH}\approx-\f{1}{3}\,\tau\,\nabla\cJ$.
The moment behavior are governed by (e.g., \cite{jinLevermore_1996})
\begin{equation}
  \pd{\cJ}{t} + \nabla\cdot\vect{\cH} = 0
  \quad\text{and}\quad
  \vect{\cH} = - \tau\,\nabla\cdot\vect{\cK}.  
  \label{eq:diffusionLimit}
\end{equation}
In the context of IMEX schemes, the above relationships imply that the relation $A\,\vec{\vect{\cH}}=-\f{1}{3}\,\tau\,\tilde{A}\,\nabla\vec{\cJ}$ should hold. 
That's
\begin{equation}
   \vect{e}_{i}^{T}A^{-1}\tilde{A}\,\vect{e} = 1, \quad i=2,\ldots,s,
\end{equation}
where $\vect{e}_{i}$ is the $i$th column of the $s\times s$ identity matrix.
\rc{It can be rewrite to $c=\tilde{c}$.}
We have proved in \cite{chu_etal_2018} that only type~ARS IMEX schemes can be diffusion accurate and constraint-preserving at the same time.

\subsection{PD-ARS schemes}

Combining the conditions we have, the 1st-order in general, high-order in streaming limit, diffusion accurate, constraint-preserving IMEX schemes we need are type~ARS IMEX schemes satisfying the following constraints:
\begin{enumerate}
    \item Consistency of the implicit coefficients:
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
    \end{equation}
    \item High-order accuracy in the streaming limit, 2nd-order for example:
    \begin{equation}
      \sum_{i=1}^{s}\tilde{w}_{i}=1
      \quad\text{and}\quad
      \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},
      \label{eq:orderConditionsEx}
    \end{equation}
    \item Diffusion accurate: \rc{$c=\tilde{c}$.}
    \begin{equation}
      \vect{e}_{i}^{T}A^{-1}\tilde{A}\,\vect{e} = 1, \quad i=2,\ldots,s.
      \label{eq:diffusionCondition}
    \end{equation}
    \item Convex-invariant:
    \begin{align}
      &a_{ii}>0, \quad c_{i0},\tilde{c}_{i0}\ge0, \quad \text{for} \quad i=2,\ldots,s, \nonumber \\
      &\text{and} \quad c_{ij},\tilde{c}_{ij}\ge0, \quad \text{for} \quad i=3,\ldots,s, \quad\text{and}\quad j=2,\ldots,i-1.  
      \label{eq:convexInvariant}
    \end{align}
    with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\Sch}=\min_{\substack{i = 2,\ldots,s \\ 
                  j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}}>0$.
    \item Less than five stages ($s\le4$) \label{cod:statges}.
    \item Globally stiffly accurate: $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$. 
\end{enumerate}  
The \ref{cod:statges} constraint is given by efficiency consideration.
We call the IMEX scheme satisfying the above conditions {PD-ARS}. (Definition 3 in \cite{chu_etal_2018}.)
We give two optimized PD-ARS schemes here: PD-ARS with SSPRK2 and  PD-ARS with SSPRK3. 
They have second- and third-order accuracy in the streaming limit, respectively.
\subsubsection{PD-ARS with SSPRK2}
Here we give the optimized PD-ARS with SSPRK2 in the standard double Butcher tableau form: explicit tableau on the left ($\tilde{A}$), implicit tableau on the right ($A$):
\begin{equation}
  \begin{array}{c | c c c}
  	0 & 0   & 0 & 0 \\
  	1 & 1   & 0 & 0 \\
  	1 & 1/2 & 1/2 & 0 \\ \hline
  	  & 1/2 & 1/2 & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1 & 0 & 1/2 & 1/2 \\ \hline
  	  & 0 & 1/2 & 1/2
  \end{array}
\end{equation}
For this scheme, $c_{\mbox{\tiny Sch}}= 1$ and only two implicit solver are needed for each time step.
\subsubsection{PD-ARS with SSPRK3}
Here we give the optimized PD-ARS with SSPRK3.
Its standard double Butcher tableau are (explicit tableau on the left, implicit tableau on the right)
\begin{equation}
  \begin{array}{c | c c c c}
  	    &     &     &     &  \\
  	 1  & 1   &     &     &  \\
  	1/2 & 1/4 & 1/4 &  \\
  	 1  & 1/6 & 1/6 & 2/3 &  \\ \hline
  	    & 1/6 & 1/6 & 2/3 &
  \end{array}
  \qquad
  \begin{array}{c | c c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1/2 & 0 & 1/4 & 1/4 \\ 
  	1 & 0 & 1/6 & 1/6 & 2/3\\\hline
  	  & 0 & 1/6 & 1/6 & 2/3
  \end{array}
\end{equation}
For this schemes, $c_{\mbox{\tiny Sch}}= 1$.
Three implicit solver are taken for each time step.
Since PD-ARS with SSPRK3 is as accurate as PD-ARS with SSPRK2 in collision involved region, see Section~\ref{se:NumericalTests} for the plots, it may not offer any advantage.
